---
title: "Use of R models for HTA submission"
subtitle: "Living document on individual country guidance"
author: "Rose Hart, Dark Peak Analytics Ltd."
date: "Last updated: `r format(Sys.Date(), format = '%d %B %Y')`"
output:
  rmarkdown::html_document:
    theme: flatly
---

<style type="text/css">
.main-container {
  max-width: 20000px;
  margin-left: 50px;
  margin-right: 50px;
  margin-bottom: 50px;
}
</style>

<p><a href="https://darkpeakanalytics.com/" target="_blank">
  <img src = "DPAlogo.jpg" 
       alt = 'logo'
       style = 'position:absolute; top:0; right:0; padding:10px;'
       width = '150px'
       heigth = '150px' >
</a></p>       



```{r, echo=FALSE}

yellowhex <- "#ebc622"
redhex <- "#bd0000"
greenhex <- "#5ec22c"

CountryVec <- c(
  "Australia",
  "Belgium",
  "Canada",
  "Czech Republic",
  "Denmark",
  "England",
  "Finland",
  "France",
  "Germany",
  "Ireland",
  "Italy",
  "New Zealand",
  "Norway",
  "Poland",
  "Portugal",
  "Scotland",
  #"Spain",
  "Sweden",
  "The Netherlands"
)

```


<style>
    .HTAtable {
        width: 100%;
    }
    .HTAtable th {
        background: #d3d3d2;
        word-wrap: break-word;
        text-align: center;
    }
    .HTAtable tr:nth-child(`r which(CountryVec == "Australia")`) td:nth-child(3) { background: `r yellowhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Belgium")`) td:nth-child(3) { background: `r yellowhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Belgium")`) td:nth-child(5) { background: `r yellowhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Canada")`) td:nth-child(3) { background: `r yellowhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Canada")`) td:nth-child(5) { background: `r greenhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Czech Republic")`) td:nth-child(5) { background: `r greenhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Denmark")`) td:nth-child(3) { background: `r redhex`;color:white; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Denmark")`) td:nth-child(5) { background: `r redhex`;color:white; }
    .HTAtable tr:nth-child(`r which(CountryVec == "England")`) td:nth-child(3) { background: `r greenhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "England")`) td:nth-child(5) { background: `r greenhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Finland")`) td:nth-child(3) { background: `r greenhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Finland")`) td:nth-child(5) { background: `r greenhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Germany")`) td:nth-child(5) { background: `r greenhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Ireland")`) td:nth-child(3) { background: `r yellowhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Ireland")`) td:nth-child(5) { background: `r greenhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Italy")`) td:nth-child(3) { background: `r redhex`;color:white; }
    .HTAtable tr:nth-child(`r which(CountryVec == "New Zealand")`) td:nth-child(3) { background: `r yellowhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Norway")`) td:nth-child(3) { background: `r greenhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Norway")`) td:nth-child(5) { background: `r greenhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Poland")`) td:nth-child(3) { background: `r yellowhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Portugal")`) td:nth-child(3) { background: `r yellowhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Portugal")`) td:nth-child(5) { background: `r greenhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Scotland")`) td:nth-child(5) { background: `r yellowhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "Sweden")`) td:nth-child(5) { background: `r greenhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "The Netherlands")`) td:nth-child(3) { background: `r greenhex`;color:black; }
    .HTAtable tr:nth-child(`r which(CountryVec == "The Netherlands")`) td:nth-child(5) { background: `r greenhex`;color:black; }
    .HTAtable, tr {
        border: 1px solid #9c9e9b;
    }
</style>


---

This document is intended to be a living document for individual country guidance on the use of R models as evidence in HTA submissions. This consists of both available guidance from individual HTA websites, with references, but also correspondence with individual organisations. The correspondence with each HTA agency has been made via the 'contact us' information provided on each website. 

**Disclaimer:** This document is a guide only; the author holds no responsibility for where information here differs from individual experiences. We encourage anyone interested in submitting to a HTA agency to follow the published guidance and reach out to the individual agencies to confirm acceptance based on your individual circumstance. 


This document is to be referenced as: Hart R, Use of R models for HTA submission, 2025, https://doi.org/10.5281/zenodo.15056439

The code for this document is hosted on GitHub at: https://github.com/dark-peak-analytics/HTA_acceptance_living_document

Notice something wrong or missing? Contact us at rhart@darkpeakanalytics.com with more information and we will try to incorporate it


```{r include=FALSE}
library(fontawesome)
library(ggplot2)
library(grid)
library(rworldmap)
library(dplyr)
library(plotly)
library(gridExtra)
``` 

---

<b><u>Key</b></u>


- `r fa(name = "circle", fill = greenhex, height = "1em", prefer_type = "solid")`: R models can be accepted. Please see individual cells for specifics surrounding acceptance
- `r fa(name = "circle", fill = redhex, height = "1em",prefer_type = "solid")`: R models cannot be accepted at this time.
- `r fa(name = "circle", fill = yellowhex, height = "1em",prefer_type = "solid")`: R models may be accepted, but guidance either does not specify R outright as an acceptable option or states that discussion will be needed on the rationale for using R before deciding whether it is appropriate



<br>

<div class="HTAtable">


| Country    | Agency                                  | Available guidance from agency website | Reference | Additional guidance from correspondence | Additional notes |
|:----------|:---------------------------------------|:--------------------------------------|:-----------|:---------------------------------------|:----------------|
| <b>Australia</b>  | Pharmaceutical Benefits Advisory Committee (PBAC) | States that TreeAge Pro and Excel 2010 are the only software that can be used without pre-arrangement<br><br> Requirements of submission: <br>• Access to the electronic copy of the evaluation <br> • Ensure that variables can be changed independently<br> • Can produce results within 'reasonable running times'| <a href = "https://pbac.pbs.gov.au/information/printable-version-of-guidelines.html" target="_blank">Link</a> <br>Version 5 (September 2016) | Awaiting reply |   |
| <b>Belgium</b> |Belgian Health Care Knowledge Centre (KCE)| The guidance states that the intention is to find a balance between transparency and practical feasibility: "Different software packages for building economic models are available (e.g. Excel, DATA/Treeage, R or WinBUGs). While no specific software packages are prohibited, it is important to ensure that the assessors are practically able to validate the model. Therefore, either a generally applied package is used (e.g. Excel) or prior consultation with the assessors about which package can be used or how the model can be delivered to allow for validation is recommended." <br><br> Requirements of submission: <br> •  Routine submission of the model is not required. However, the assessor can always ask for a copy of the model  | <a href = "https://kce.fgov.be/en/publications/all-reports/belgian-guidelines-for-economic-evaluations-and-budget-impact-analyses-third-edition" target="_blank">Link</a> <br> (May 2025) | No submissions in R have yet been made in Belgium | |
| <b>Canada</b> |Canada's Drug Agency (CDA)| States that the economic model must be programmed in Excel but the sponsor must contact CDA-AMC in advance if considering alternative program software to ensure that it is acceptable and whether additional requirements will apply.<br><br> Requirements of submission: <br> • The model must be able to function in a stand-alone environment that does not require access to a web-based platform. <br> • The sponsor must provide the model in its entirety and be able to fully execute the model based on modifications to parameters of interest <br> • CDA-AMC must be able to vary individual parameters, view the calculations, and run the model to generate results | <a href = "https://www.cda-amc.ca/procedures-reimbursement-reviews" target="_blank">Link</a> <br> (September 2024) |• Not yet received submission fully in R, but have received submissions where specific functionality is programmed in R (uncommon: around 5 instances in last 5 years)<br> • Would accept R, but would expect there to be strong justification<br> • Any submitted code must be clearly annotated and with minimal use of packages that need to be installed from CRAN <br> • Would need to adhere to all specified guidelines (e.g.  requires inclusion of Markov trace, ability to vary individual parameters and have locally executable model) <br>(March 2025)| |
| <b>Czech Republic</b> |CzechHTA| Software preference not specified | Not available| The Czech authority have received models in R although it is uncommon. They have also developed models internally in R and have based their recent analysis of HPV catch-up vaccination on this platform. <br><br> The key thing that they look for is good documentation, especially surrounding the localisation of inputs and model structure (March 2025)| |
| <b>Denmark</b> |Danish Medicines Council (DMC)| “The health economic analysis and budget impact analysis must be submitted in Excel format.” | <a href = "https://medicinraadet-dk.b-cdn.net/media/huyhssl2/the-danish-medicines-councils-process-guide-for-assessing-new-pharmaceuticals-version-1-2.pdf" target="_blank">Link</a> <br> (June 2022) | R is not currently accepted (March 2025) | |
| <b>England</b> |National Institute for Health and Care Excellence (NICE)|  NICE accepts fully executable economic models using standard software, that is, Excel, DATA/Treeage, R or WinBUGs <br><br>  Requirements of submission: <br> • NICE must have access to the programming code of a fully executable model <br> • No specifications on package preference or code guidelines| <a href = "https://www.nice.org.uk/process/pmg36/resources/nice-health-technology-evaluations-the-manual-pdf-72286779244741" target="_blank">Link</a> <br> (October 2023) | Confirmed that R models have been submitted and accepted. <br><br> There are no plans for developing guidelines/preferences on coding structure and packages to be used. (April 2025)| |
| <b>Finland</b> |Pharmaceuticals Pricing Board| States that recommended softwares for executing the model are Microsoft Excel and R. <br><br>It must be possible to view and edit the calculation formulas and initial data in the electronic document. The model needs to facilitate the addition of new scenarios and the model must not be password protected. | <a href = "https://www.hila.fi/content/uploads/2024/06/Instructions_TTS_2024_190624.pdf" target="_blank">Link</a> <br> (June 2024) | The Pharmaceuticals Pricing Board has not yet received any R model as part of an application submission. R packages published on CRAN are recommended. Stated that all instructions for health economic evaluations currently available apply to R models. <br><br> The model should be attached and the Pharmaceuticals Pricing Board should be able to view and edit the model code and data, including creating new scenarios. l it is also important to have comprehensive comments in the code to ensure smooth assessment process. (March 2025) | |
| <b>France</b> |Haute Autorité de Santé (HAS)| Software preference not specified | <a href = "https://www.has-sante.fr/upload/docs/application/pdf/2014-03/pricing_reimbursement_of_drugs_and_hta_policies_in_france.pdf" target="_blank">Link</a> <br> (March 2014) | Awaiting reply | |
| <b>Germany</b> |Institute for Quality and efficiency in Health Care (IQWiG)| Software preference not specified in English General methods guidance| <a href = "https://www.iqwig.de/en/about-us/methods/methods-paper/" target="_blank">Link</a> <br>Version 7 (September 2023) | IQWiG have stated that R is a recognized program and models developed with it will be accepted. They have confirmed that they have not yet received any models in R and that they have no current guidelines on how a model in R would need to be developed for submission. (February 2025)| |
| <b>Ireland</b> |National Centre for Pharmacoeconomics (NCPE)| States that Microsoft Excel 2021 (or earlier versions) is the preferred software for NCPE submissions.  Contact the NCPE in advance of submission if alternative software packages are to be considered for submission <br><br>Requirements of submission: <br> •  A Technical Specification Document is required to be submitted alongside the model for how users should use and adapt the model <br>  • The model must be fully executable <br>  • Specifications of outputs are detailed in guidance and would need to be adhered to regardless of software used | <a href = "https://www.ncpe.ie/submission-process/submission-templates/format-of-full-submissions/" target="_blank">Link</a> <br>Version 3.1 (March 2023) | The NCPE Will accept R with prior notification. No submissions have been made in R to date <br><br>All code would need to be provided. There is currently no guidance on preferred packages or code guidelines (March 2025) | |
| <b>Italy</b> |Italian Medicines Agency (AIFA)| “The data and models used must be made available at the same time in Excel or TreeAge format” (translated)  | <a href = "https://www.aifa.gov.it/en/linea-guida-capitolo-9" target="_blank">Link</a> <br>(January 2021) | Awaiting reply | |
|<b>New Zealand</b> |PHARMAC| PHARMAC’s preferred software packages are TreeAge and Microsoft Excel. Models provided in other software packages will not be assessed unless by prior agreement <br><br> Requirements of submission: <br> • The model must be stand-alone, where PHARMAC can access all model materials for review and amend any relevant inputs <br> • Economic models should not be unnecessarily complex, and should always be transparent, well described and reproducible | <a href = "https://www.pharmac.govt.nz/medicine-funding-and-supply/the-funding-process/policies-manuals-and-processes/economic-analysis/prescription-for-pharmacoeconomic-analysis-methods-for-cost-utility-analysis/2-health-economic-analysis-at-pharmac" target="_blank">Link</a> <br>(accessed March 2025)| Awaiting reply | |
| <b>Norway</b> |Norwegian Medicines Agency (NoMA)| States that normally it would be useful if the model is designed using Excel, but other alternatives can also be acceptable such as, TreeAge and R. <br><br>Requirements of submission<br>• NoMA must be able to change all relevant variables and parameters in the model <br> • The model must not be locked, time limited, password-protected unless the password is made available, or have any hidden elements that are not described or cannot easily be changed. <br> • The model must be fully executable and not in any non-transparent programmes and/or programming languages | <a href = "https://www.dmp.no/globalassets/documents/english/public-funding-and-pricing/documentation-for-sta/guidelines-20.05.2020.pdf" target="_blank">Link</a> <br> (May 2020) | Awaiting reply on additional information request | |
| <b>Poland</b> |Agency for Health Technology Assessment and Tariff System (AOTM)| “Except for justified situations, the models should be developed using commonly available tools - Software like MS Excel, TreeAge is often used.” (translated) | <a href = "https://www.aotm.gov.pl/en/guidelines/" target="_blank">Link</a> <br> Version 3 (August 2016) | Awaiting reply | |
| <b>Portugal</b> |Health Technology Assessment Committee (Infarmed)| The guidance states that a model software needs to be appropriate for the cost effectiveness model <br><br>Requirements of submission<br>• No specific guidance on model deployment <br>• There is to be a specific focus on drivers and outcomes of paramenter uncertainty over time| <a href = "https://www.infarmed.pt/web/infarmed/entidades/medicamentos-uso-humano/avaliacao-tecnologias-saude/avaliacao-terapeutica-e-economica2018/-i-guidelines-/i-" target="_blank">Link</a> <br> Version 1 (December 2019) | Infarmed have confirmed that they will accept submissions in R code. However noted that specific particularities of the economic model may need to be replicated in Excel if the methodology is unclear. <br><br> There is no current guidance on specific packages or coding structure. (March 2025)| |
| <b>Scotland</b> |Scottish Medicines Consortium (SMC)| Software preference not specified <br><br> Requires an executable electronic copy of the economic model with the submission, with full access to the programming code</b>| <a href = "https://scottishmedicines.org.uk/making-a-submission/" target="_blank">Link</a> <br> Guidance on NPAF (November 2024) | SMC have not accepted, nor had the request to accept any models in R. There is concern that they would not be resources to make all the required checks within their very rapid assessment timelines <br><br> They have confirmed that their ability to accept would depend on the specifics of the case and the rationale for using R would need to be considered before deciding whether it was appropriate <br><br>There is no current guidance on specific packages or coding structure. (March 2025) | |
| <b>Sweden</b> |Dental and Pharmaceutical Benefits Agency (TLV)| Software preference not specified| <a href = "https://www.tlv.se/lakemedelsforetag/halsoekonomiska-bedomningar-och-rapporter-kliniklakemedel.html" target="_blank">Link</a> <br> (February 2025) | The TLV requirement for models is that they are in electronic form - they confirmed R would therefore be accepted. They stated that the model must be fully executable and completely available to review the code and amend the inputs. <br><br>There is no current guidance on specific packages or coding structure. (March 2025)| |
|<b>The Netherlands</b> |National Health Care Institute (ZiN)| ZIN has provided a guideline specifically for the building of models in R <br><br>Details of R model submission: <br> •  CE-models in R are required to include at least all functionalities that would otherwise be included in Excel (including the potential for end-to-end functionality with necessary data)<br>• A recommended folder structure and code structure is provided<br> • The code needs to be able to function stand-alone <br> • Coding conventions and commenting guidance is described <br> • ZIN requires the model to only use CRAN packages that are listed in the guidance <br> • Unit testing is required | <a href = "https://english.zorginstituutnederland.nl/publications/publications/2022/12/15/guideline-for-building-cost-effectiveness-models-in-r" target="_blank">Link</a> <br> (December 2022) | Awaiting reply on additional information request| |


</div>



